---
title: "cpsbanalysis"
author: "mw"
date: "January 26, 2018"
output: html_document
runtime: shiny
---

```{r input, echo = FALSE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width=10, 
                      fig.height=5,
                      echo=FALSE,
                      warning=FALSE)

require(ggplot2)
require(psych)
require(dplyr)
require(Hmisc)

colMax <- function(i) sapply(i, max, na.rm = TRUE)

files <- list.files(path=".", pattern="*.csv")

data <- lapply(files, read.csv, sep=",", header=TRUE, stringsAsFactor = FALSE)
data <- do.call("rbind", data)

# Clean up the columns
names(data)[names(data) == "X"] <- "line"

names(data)[names(data) == "X.Warn"] <- "warn"
data$warn <- as.factor(data$warn)

data$OverallPlace <- as.integer(data$OverallPlace)

data$Division <- as.factor(data$Division)

data$OverallPlace <- as.integer(data$OverallPlace)

data$PlayTime <- as.difftime(data$PlayTime, format = "%H:%M", units = "mins")

# [TODO] Find a way to generalize the image names
totals <- colMax(data[names(data)[grep(".Found$", names(data))]] +
         data[names(data)[grep(".Remaining$", names(data))]])
names(totals) <- gsub(x = names(totals), pattern = "Found", replacement = "Total") 

# Formats all of the "<image>.Time" columns into minutes
data[names(data)[grep(".Time", names(data))]] <- sapply(data[names(data)[grep(".Time", names(data))]], as.difftime, format = "%H:%M", units = "mins")

# Coerce all of the "<image>.Warn" vectors to factors
data[names(data)[grep(".Warn", names(data))]] <- lapply(data[names(data)[grep(".Warn", names(data))]], as.factor)

data$State <- as.factor(data$State)
data$StatePlace <- as.integer(data$StatePlace)
data$TeamName <- as.factor(data$TeamName)
data$TeamNumber <- as.factor(data$TeamNumber)
data$Tier <- as.factor(data$Tier)

# Remove unused columns (i.e. in this particular analysis)
data$Tier <- NULL
data$Division <- NULL

```

# Analysis of Cyber Patriot State Round scores

## INTRODUCTION
The National Youth Cyber Defense Competition is the centerpiece of the Cyberpatriot program. The competition puts teams of high school and middle school students in the position of newly hired IT professionals tasked with managing the network of a small company. In the rounds of competition, teams are given a set of virtual images that represent operating systems and are tasked with finding cybersecurity vulnerabilities within the images and hardening the system while maintaining critical services. Teams compete during a six hour period for the top placement within their state and region. The state competiion took place on January 19-21, 2018. The top half of teams in each state progresses to the state semi-final round (February 9-11), where the top 3 teams advance to the National Finals Competition in Baltimore, MD. The top national teams can earn national recognition and scholarship money.

Evans Middle School started competing this year with two teams of 5 students each. The Tax Collectors (10-4784) and Users Unknown (10-4785). The following analysis compares both teams' performance among the local, state, and national results from the state roudn in order to develop  strategic insights. Such insights can influence the coach and mentors to focus training materials in a way that increases the teams' competitiveness.

## DESCRIPTION 

During each round, the unofficial score are delivered in near-real time via the competition website (http://scoreboard.uscyberpatriot.org/). In addition to displaying key competition data, such as current score and elapsed time, a secondary link provides additional data, such as a break down of artifacts found, artifacts remaining, and score per image as well as any details on any penalties.

After the state round, 

``````{r summary, echo = FALSE, message = FALSE, warning = FALSE}
  s

```

## COMPETITION:
* Overall completion (number and ratio) for each of the three images.
* Comparison of the state, top 25, and bottom 25 among our results.
* Which image has the lowest score among the best performers?

## OVERALL:
* How are image scores distributed among common total scores?
* How do penalties influence the image/total scores?
* How much time do team spend on images with common scores?



This R Markdown document is made interactive using Shiny. Unlike the more traditional workflow of creating static reports, you can now create documents that allow your readers to change the assumptions underlying your analysis and see the results immediately. 

To learn more, see [Interactive Documents](http://rmarkdown.rstudio.com/authoring_shiny.html).





## Inputs and Outputs

You can embed Shiny inputs and outputs in your document. Outputs are automatically updated whenever inputs change.  This demonstrates how a standard R plot can be made interactive by wrapping it in the Shiny `renderPlot` function. The `selectInput` and `sliderInput` functions create the input widgets used to drive the plot.

```{r eruptions, echo=FALSE}
inputPanel(
  selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),
  
  sliderInput("bw_adjust", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2)
)

renderPlot({
  hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Duration (minutes)", main = "Geyser eruption duration")
  
  dens <- density(faithful$eruptions, adjust = input$bw_adjust)
  lines(dens, col = "blue")
})
```

## Embedded Application

It's also possible to embed an entire Shiny application within an R Markdown document using the `shinyAppDir` function. This example embeds a Shiny application located in another directory:

```{r tabsets, echo=FALSE}
shinyAppDir(
  system.file("examples/06_tabsets", package = "shiny"),
  options = list(
    width = "100%", height = 550
  )
)
```

Note the use of the `height` parameter to determine how much vertical space the embedded application should occupy.

You can also use the `shinyApp` function to define an application inline rather then in an external directory.

In all of R code chunks above the `echo = FALSE` attribute is used. This is to prevent the R code within the chunk from rendering in the document alongside the Shiny components.



