---
title: "Analysis of Cyber Patriot 11 First Round scores"
author: "mw"
date: "November 28, 2018"
output: html_document
runtime: shiny
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
  
  knitr::opts_chunk$set(fig.width=10, fig.height=6, echo=FALSE, warning=FALSE)
  
  require(ggplot2)
  require(psych)
  require(dplyr)
  require(scales)
  require(grid)
  require(gridExtra)
  require(reshape2)

  
  imagenames <- c("Windows Server 2016", "Windows 10")
  teams <- c("11-3500", "11-3502", "11-3503")
  
  colMax <- function(i) sapply(i, max, na.rm = TRUE)
  
  files <- list.files(path=".", pattern="^CP.*csv")
  
  data <- lapply(files, read.csv, sep=",", header=TRUE, stringsAsFactor = FALSE)
  data <- do.call("rbind", data)
  
  # Clean up the columns
  names(data)[names(data) == "X"] <- "line"
  
  names(data)[names(data) == "X.Warn"] <- "warn"
  data$warn <- as.factor(data$warn)
  data$OverallPlace <- as.integer(data$OverallPlace)
  data$Division <- as.factor(data$Division)
  data$OverallPlace <- as.integer(data$OverallPlace)
  data$PlayTime <- as.difftime(data$PlayTime, format = "%H:%M", units = "mins")
  
  # [TODO] Clean up the image name suffixes
  names(data) <- gsub(x = names(data), pattern = "_cpxi_r1_ms.", replacement = ".")
  names(data) <- gsub(x = names(data), pattern = "_cpxi_r1_hsms.", replacement = ".")
  
  # Summarize the images and total artifacts
  totals <- colMax(data[names(data)[grep(".Found$", names(data))]] +
           data[names(data)[grep(".Remaining$", names(data))]])
  names(totals) <- gsub(x = names(totals), pattern = "Found", replacement = "Total") 
  names(totals) <- unlist(lapply(strsplit(names(totals), "_"), `[[`, 1))
  
  # Add Known team names from lookup file
  lookup <- read.csv("lookup.csv", sep=",", header = TRUE, stringsAsFactor = TRUE)
  data <- merge(lookup, data, by = "TeamNumber", all=TRUE, incomparables = '')
  data$TeamName <- NULL
  names(data)[names(data) == 'Name'] <- 'TeamName'
  data$TeamName <- as.factor(data$TeamName)
  excludes <- data[is.na(data$line),]
  data <- data[!is.na(data$line),]
  

  # Formats all of the "<image>.Time" columns into minutes
  data[names(data)[grep(".Time", names(data))]] <- sapply(data[names(data)[grep(".Time", names(data))]], as.difftime, format = "%H:%M", units = "mins")
  
  # Coerce all of the "<image>.Warn" vectors to factors
  data[names(data)[grep(".Warn", names(data))]] <- lapply(data[names(data)[grep(".Warn", names(data))]], as.factor)
  
  # Add the completion ratios to each team's scores
  # [TODO] Change the field names to match the competition images
  data$Server2016.Completion <- data$Server2016.Found/totals[1]
  data$Windows10.Completion <- data$Windows10.Found/totals[2]
  
  data$State <- as.factor(data$State)
  data$StatePlace <- as.integer(data$StatePlace)
  data$TeamNumber <- as.factor(data$TeamNumber)
  data$Tier <- as.factor(data$Tier)
  
  row.names(data) <- NULL
  
  
  # Remove unused columns (i.e. in this particular analysis)
  data$Tier <- NULL
  data$Division <- NULL
  data$X.1 <- NULL
  
  data <- data[order(data$OverallPlace),]
  top50 <- droplevels(data[data$OverallPlace >= 1 & data$OverallPlace <= 25,])
  ga <- droplevels(data[data$State == "GA",])
  cc <- droplevels(data[!is.na(data$TeamName),])
  top50 <- top50[order(top50$OverallPlace),]
  ga <- ga[order(ga$StatePlace),]
  cc <- ga[order(cc$StatePlace),]
  
  nteams <- nrow(data)
  myteams <- droplevels(data[data$TeamNumber %in% teams,])
#  data$CurrentScore[is.na(data$CurrentScore)] <- 0

```

# INTRODUCTION
The National Youth Cyber Defense Competition is the centerpiece of the Cyberpatriot program. 
Teams compete by finding cybersecurity vulnerabilities on a set of virtual images, then fixing them while maintaining critical services. 
Teams compete during a six hour period for the top placement within their state and region. 
The first round of the 2018 competiion season took place on November 2-4, 2018. 

The second round takes place on December 7-9, 2018 and the State round on January 11-13, 2019. 
The top half of teams in each state progresses to the state semi-final round (February 1-3, 2019), then the top 3 teams advance to the National Finals Competition in Baltimore, MD on April 7-11, 2019. 

<!-- Team,Name -->
<!-- 11-3500,Tax Collectors V2 -->
<!-- 11-3502,Bits and Bytes -->
<!-- 11-3503,Digital Decoders -->

Evans Middle School started competing last year in CP-10 with two five-student teams. This second year of competing expanded to three teams; two with six students, and one with five students. 
The Evans Middle School teams are the *````Tax Collectors V2````* (```11-3500```), *```Bits & Bytes```* (```11-3502```), and *```Digital Decoders```* (```11-3503```). 
Both teams competed against a total of ```r nrow(data)``` teams in the nation, and ```r nrow(ga)``` in Georgia--```r nrow(cc)``` of which are from other Columbia County schools. 

The following analysis compares the Evans Middle School teams' performance among the national, state, and local results from the first round in order to develop strategic insights for subsequent rounds. 
Such insights can influence the coach and mentors to focus training materials in a way that increases the teams' competitiveness.

# COMPETITION DESCRIPTION 

During each round, the unofficial score are delivered in near-real time via the competition website (```http://scoreboard.uscyberpatriot.org/```). 
The main page displays a table of key competition data, such as current score and elapsed time. 
Each row on the table contains a secondary link that provides supplementary data, such as a the number of artifacts found, artifacts remaining, and score per image as well as details on any penalties.

The first round featured ```r length(names(totals))``` images: ```r imagenames```. 
The table below shows the total number of scored artifacts available on each image during this round.

```{r imagesummary, tidy=TRUE}
  print(totals)
```

# SCORING ANALYSIS 

The overall results from the Evans Middle School teams are in the table below.
_Note that the units for ```PlayTime``` fields are elapsed minutes since the image was initialized._ 
_Teams have six continuous hours (i.e., 360 minutes) and must officially close their image before the deadline to avoid a penalty._

## Team Results

```{r myteamresults}
  vars <- c("TeamName", "StatePlace", "OverallPlace", "CurrentScore", "PlayTime")
  print(myteams[vars], row.names = FALSE)
```

The following tables summarize the team results for each of the images. 
More detailed analysis is included in the _Completion Analysis_ section, later in this document.

### Windows Server 2016 (`r totals[["Server2016.Total"]]` possible artifacts)

```{r myteam2016}
  vars <- c("TeamName", "Server2016.Found", "Server2016.Completion", "Server2016.Score", "Server2016.Time")
  disp <- myteams[vars]
  names(disp) <- c("Name", "Found", "Complete(%)", "Score", "Time(min)")
  disp$`Complete(%)` <- percent(disp$`Complete(%)`)
  print(disp, row.names = FALSE)
```

### Windows 10 (`r totals[["Windows10.Total"]]` possible artifacts)
```{r myteamwindows10}
  vars <- c("TeamName", "Windows10.Found", "Windows10.Completion", "Windows10.Score", "Windows10.Time")
  disp <- myteams[vars]
  names(disp) <- c("Name", "Found", "Complete(%)", "Score", "Time(min)")
  disp$`Complete(%)` <- percent(disp$`Complete(%)`)
  print(disp, row.names = FALSE)
```


## Competition Results

The analysis in this section presents the overall results of the first round. 
Then, the results of the Evans Middle School teams is compared against the national, state results, and the local community results.

### Top `n` Teams
```{r topnsummary}
  # Extract the subsets of interest
  numericInput("rows", "How many Teams?", 25)
  
  renderTable({
    vars <- c("OverallPlace", "TeamNumber", "State", "StatePlace", "CurrentScore")
    head(data[vars], input$rows)
  })
```

### Scores from Georgia schools
Team listed in the table below are registered but did not compete in the first round.
```{r excludes}
  vars <- c("TeamNumber", "TeamName")
  print(excludes[vars], row.names = FALSE)
```
_NOTE: Teams with a `TeamName` are the Columbia County middle schools with Cyberpatriot teams._ 
```{r statesummary}
  renderTable({
    vars <- c("StatePlace", "TeamNumber", "TeamName", "OverallPlace", "CurrentScore")
    head(ga[vars], nrow(ga))  
  }, row.names = FALSE)
  
```

## Score Comparision

The plots below summarizes the unofficial scores by state. 
The numbers along the bottom of the x-axis note the number of teams in each state. 
The horizontal dashed line represents the national median score (`r median(data$CurrentScore)`). 
Blue highlights represent Georgia on the left plot, and the Columbia County Middle Schools on the right plot. 
The Evans Middle School teams are highlighted in red on the right plot.

Overall, the Evans Middle School teams are in the upper quartile (i.e., top 25%) for the state of Georgia, 


```{r score}
  give.n <- function(x){
  return(c(y = -10, label = length(x))) 
  }

  p1 <- ggplot(data, aes(State, CurrentScore)) +
    geom_boxplot() +
    ggtitle("Overall Scores by State") + labs(x = "State", y = "Score") +
    stat_summary(fun.data = give.n, geom = "text", position = position_dodge(width = 1), fun.y = median) +
    geom_boxplot(data=ga, aes(State, CurrentScore),fill="blue") +
    geom_line() + geom_hline(yintercept = median(data$CurrentScore), color="blue", lty=2)
    
  p2 <- ggplot(data=ga, aes(State, CurrentScore)) +
      geom_boxplot(outlier.size=0) +
      geom_jitter(data=data[data$TeamName != "",], aes(State, CurrentScore),
             position=position_jitter(width=0.1,height=0.1),
             alpha=0.6,
             size=3,
             show.legend = FALSE,
             color = "blue") +
      geom_jitter(data=myteams, aes(State, CurrentScore),
               position=position_jitter(width=0.1,height=0.1),
               alpha=0.6,
               size=3,
               show.legend = FALSE,
               color = "red") +
    xlab("Georgia Teams") +
    ylab("Score") +
    ggtitle("GA Teams")
    
    grid.arrange(p1, p2, ncol = 2, top = "Score Analysis", widths = c(5, 1))
    
    dens <- data %>% group_by(State) %>% summarise(nteams = n(), medianscore = median(CurrentScore))
```

### Score Density

Overall placement is determined primarily by overall scores (higher is better).
Numerous teams end up with the same overall score, so time (faster is better) is used to further categorize team placement.
The plots below summarize the Overall Scores and Play Times to provide descriptive context of both variables for the remaining analysis.

```{r densitystats}
  tm <- data %>% select(OverallPlace, CurrentScore, PlayTime)
  d <- describeBy(tm[c(-1)])
  print(d[,c(-1,-6,-7)], row.names = FALSE)
  
  d <- melt(tm, id.vars="OverallPlace")
  p1 <- ggplot(d, aes(variable, value)) + geom_boxplot() +
    xlab("Variable") +
    ylab("Value") +
    ggtitle("Summary of Score and Time Variables")
  p1
  
```

The plots below illustrate the density of scores and time.
The red bars represent the Evans Middle School teams.
The red bars represent the State of Georgia teams.
The gray bar represent the teams from the rest of the nation or world.
The vertical dotted line represents the median.
Note that while the Georgia teams are spread on either side of the median, the Evans Middle School teams were consistently above the median.


```{r scoretimedensity}
  p1 <- ggplot(data, aes(CurrentScore)) +
      geom_histogram(binwidth = 1) +
      geom_histogram(data=ga, fill="blue", binwidth = 1) + 
      geom_histogram(data=myteams, fill="red", binwidth = 1) + 
      scale_x_continuous(breaks = seq(0,200,10)) +
      scale_y_continuous(breaks=seq(0,25,2)) +
      geom_vline(xintercept = median(data$CurrentScore, na.rm = TRUE), lty=2) +
      ggtitle("Scoring Density") + labs(x = "Overall Score", y = "Number of Teams")
  
  p1
  
  p2 <- ggplot(data, aes(PlayTime)) +
      geom_histogram(binwidth = 10) + 
      geom_histogram(data=ga, fill="blue", binwidth = 10) + 
      geom_histogram(data=myteams, fill="red", binwidth = 10) + 
      geom_vline(xintercept = median(data$PlayTime, na.rm = TRUE), lty=2) +
      ggtitle("Play Time Density") + labs(x = "Play Time (minutes)", y = "Number of Teams")
  
  p2

```

### [DRAFT] Score Density Compared to Time

NOTE: This section is still being developed.

```{r scorevstime}
  p1 <- ggplot(tm, aes(x=CurrentScore, y=PlayTime)) + 
    geom_point(shape = 1) + 
    geom_smooth(method = lm)
  
  p1

```

### Impact of State Team Density on Overall Scores

A total of ```r nrow(dens)``` states and countries participated in this round. 
The table of diagnotic statistics below shows that each state has a large range of the number of teams as well and overall scores. 
It may seem intuitive that states with more teams lead to better overall scores, 
However, a comparison of the number of teams in each state against the state's median score suggests that there is very little statistical relationship between the two.
The figures below summarize the descriptive statistics and density of teams per state/country.

```{r teamdensitystats}
  d <- describeBy(dens[c(-1)])
  print(d[,c(-1,-6,-7)], row.names = FALSE)
```


```{r stateteams}
  stteams <- data.frame(table(data$State))
  names(stteams) <- c("State", "Teams")
  
  p1 <- tableGrob(head(stteams[order(-stteams$Teams),],10), rows = NULL)
  
  p2 <- ggplot(stteams, aes(x=reorder(State, -Teams), Teams)) +
      geom_bar(stat = "identity") +
      labs(x = "State", y = "Teams")

  grid.arrange(p1, p2, ncol = 2, top = "Team Density by State", widths = c(1, 4))
  
```

The plots below illustrates the states with the highest overall scores tend to have very few teams, while the states with the most teams trend along overall median. 

The boxplot on the left provides the context of overall team density.
The median density for a state is ```r quantile(dens$nteams)[["50%"]]``` teams, with an IQR of ```r quantile(dens$nteams)[["75%"]] - quantile(dens$nteams)[["25%"]]```.
The outliers are plotted vertically.
Georgia, with ```r dens[dens$State=="GA",]$nteams``` teams is one of the outliers.

In contrast, the scatterplot on the right shows each state/country horizontally against their median score.
The points in red show the lowest and highest density states in relation to Georgia.
The blue dashed line represents the overall median of the state median scores (```r quantile(dens$medianscore)[["50%"]]```) with and IQR of ```r quantile(dens$medianscore)[["75%"]] - quantile(dens$medianscore)[["25%"]]```; the IQR is omitted from the scatterplot for clarity.
A small, negative correlation coefficient (```r round(cor(dens$nteams, dens$medianscore), 4)```) suggests that more teams per state do not necessarily translate into higher median scores.
The median score for Georgia (```r dens[dens$State=="GA",]$medianscore```) is slightly higher than the overall median score, and within the IQR. 
Therefore, the state of Georgia is competitive but not necessarily due to the higher team density.


```{r teamdensity}
  p1 <- ggplot(dens, aes(x = "", y = nteams)) + 
    geom_boxplot() + 
    labs(x = "", y = "Overall Team Density per State")

  p2 <- ggplot(dens, aes(x = nteams, y = medianscore)) + 
  geom_point(aes(colour = ifelse((dens$State == "GA" | dens$nteams == max(dens$nteams) | dens$medianscore == max(dens$medianscore)), "black", "red")))+
  geom_text(aes(label = ifelse((dens$State == "GA" | dens$nteams == max(dens$nteams) | dens$medianscore == max(dens$medianscore)), paste0(as.character(State), "(", as.character(nteams),")"), "")), hjust = 1, vjust = -0.75, size = 3)+
  theme(legend.position = "none") +
  annotate("text", label = paste0("cor = ", round(cor(dens$nteams, dens$medianscore), 4)), x = 100, y = 0) +
  geom_hline(yintercept = median(data$CurrentScore), color= "blue", lty=2) + 
  labs(x = "Number of teams", y = "Median Score")
  
  grid.arrange(p1, p2, ncol = 2, top = "Team Density Analysis", widths = c(1, 5))
```
_NOTE: Each point represents a State's median score in relation to the number of teams in that state. Therefore, the horizontal blue dashed line represents the median of the overall scores, and not the median of the points represented in the plot (i.e., the median of medians)._

# IMAGE ANALYSIS

## Image Scoring

```{r completionsetup}
  d <- data.frame(table(data$ScoredImages))
  names(d) <- c("Scored Images", "Teams")
```

The Evans Middle School teams attempted both of the competition images, consistent with ```r percent(d$Teams[2]/nrow(data))``` of the teams in the competition. 
In contrast, ```r d$Teams[1] + d$Teams[2]``` teams or ```r percent((d$Teams[1] + d$Teams[2])/nrow(data))``` (including ```r nrow(droplevels(ga[ga$ScoredImages < 2,]))``` in Georgia) attempted less than two images. 
The teams that did not attempt both images ended up placing between ```r min(data[data$ScoredImages < 2,]$OverallPlace)``` and ```r max(data[data$ScoredImages < 2,]$OverallPlace)```.

The plot below shows a large variation among the scores with teams that attempted both images. 
However, the scores of the teams that did not attempt both images are not competitive.

```{r attempts}

  ggplot(data, aes(x=ScoredImages, y = CurrentScore)) + 
  geom_boxplot(aes(group=ScoredImages)) + 
  geom_jitter(color = "grey50") +
      geom_jitter(data=data[data$TeamName != "",], aes(ScoredImages, CurrentScore),
             position=position_jitter(width=0.05,height=0.1),
             alpha=0.6,
             size=3,
             show.legend = FALSE,
             color = "blue") +
      geom_jitter(data=myteams, aes(ScoredImages, CurrentScore),
               position=position_jitter(width=0.05,height=0.1),
               alpha=0.6,
               size=3,
               show.legend = FALSE,
               color = "red") +
      ggtitle("Scores per Images") + labs(x = "Attempted Images", y = "Score")

```


## Image Completion Rates

The plots below show the ratio of artifacts found compared to number of total artifacts on each image. 
Note that a ratio is used (i.e., instead of number of artifacts found) in order to compare the two images of different numbers of available artifacts. 
In this case, both images have the same number of artifacts, so either approach would be sufficient for comparison.

Both of the Evans Middle School teams completed more on each image than the nationwide median. More specifically, both teams had a significantly higher completion ratio on the Windows 10 images. 
The ```Bits and Bytes``` had a lower completion ratio on the Windows Server 2016 image, while ```Digital Decoders``` and ```Tax Collectors v2``` had the same completion ratio. 
All three teams had the same completion ratio on the Windows 10 image.

```{r completion}
  d <- data[,c(1, 26:27)]
  d <- melt(data = d, id.vars = "TeamNumber", measure.vars = c("Server2016.Completion", "Windows10.Completion" ))
  d$variable <- gsub(".Completion", "", d$variable)
  names(d) <- c("Teams", "Image", "Complete")
  
  db <- describeBy(d[3], group = d$Image, mat = TRUE, digits = 2)
  
  print(db[,c(-1,-3,-8,-9)], row.names = FALSE)
  
  ggplot(d, aes(x=Image, y = Complete)) +
      geom_boxplot() +
    
      geom_jitter(data=d[d$Teams %in% ga$TeamNumber,], aes(Image, Complete),
               #position=position_jitter(width=0.1,height=0.1),
               alpha=0.6,
               size=3,
               show.legend = FALSE,
               color = "blue") +

      geom_jitter(data=d[d$Teams %in% teams,], aes(Image, Complete),
         #position=position_jitter(width=0.1,height=0.1),
         alpha=0.6,
         size=3,
         show.legend = FALSE,
         color = "red") +
    
      ggtitle("Overall Completion Rates per Image") + labs(x = "State", y = "Score") +
      ylab("% Complete")
```

## Image Artifacts

The plots below count the number of teams that found the same number of artifacts on each image.
The results are consistent with the previosuly stated observations.

```{r imagescore}
  g1 <- ggplot(data, aes(Server2016.Found)) + geom_bar() +
  geom_bar(data=ga, aes(Server2016.Found),fill="blue") +
  geom_bar(data=myteams, aes(Server2016.Found),fill="red", width = 1)

# aes(fill = ifelse(data[data$TeamNumber %in% teams,]$Server2016.Found, "grey50", "red"))

  g2 <- ggplot(data, aes(Windows10.Found)) + geom_bar() + 
    geom_bar(data=ga, aes(Windows10.Found),fill="blue") +
    geom_bar(data=myteams, aes(Windows10.Found),fill="red", width = 1)
    
  grid.arrange(g1, g2, ncol = 2, top = "Frequency of Finds per Image")



```


# CONCLUSIONS
* Compared to the National and State teams, both of the Evans Middle School teams are very competitve on the Windows Server image, but could have been more competitive on the Windows 10 image.
* The number of teams that compete in a given state does not necessarily have a positive effect on overall scores. In fact, the states with between ```7``` and ```12``` teams appear the most competitive.
* Teams that do not attempt both images are not competitive, though attempting both images does not necessarily guarantee high placement.
* [DRAFT] Is there an advantage to finishing earlier?