---
title: "Analysis of Cyber Patriot 12 Round 2"
author: "mw"
date: "November 18, 2019"
output: html_document
runtime: shiny
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
  
  knitr::opts_chunk$set(fig.width=10, fig.height=6, echo=FALSE, warning=FALSE)
  
  require(ggplot2)
  require(psych)
  require(dplyr)
  require(scales)
  require(grid)
  require(gridExtra)
  require(reshape2)

  
  imagenames <- c("Server 2016", "Windows 10", "Ubuntu 14")
  teams <- c("12-1306","12-1307","12-1308","12-1309")
  
  colMax <- function(i) sapply(i, max, na.rm = TRUE)
  
  files <- list.files(path=".", pattern="^CP-XII Round 2.*csv")
  
  data <- lapply(files, read.csv, sep=",", header=TRUE, stringsAsFactor = FALSE)
  data <- do.call("rbind", data)
  
  # Clean up the columns
  names(data)[names(data) == "X"] <- "line"
  names(data)[names(data) == "X.Warn"] <- "warn"
  data$warn <- as.factor(data$warn)
  data$OverallPlace <- as.integer(data$OverallPlace)
  data$Division <- as.factor(data$Division)
  data$OverallPlace <- as.integer(data$OverallPlace)
  data$PlayTime <- as.difftime(data$PlayTime, format = "%H:%M", units = "mins")
  
  # [TODO] Summarize the number of teams that attempted each image
  
  # [TODO] Clean up the image name suffixes
  names(data) <- gsub(x = names(data), pattern = "_cpxii_r2_hsms.", replacement = ".")
  names(data) <- gsub(x = names(data), pattern = "_cpxii_r2_ms.", replacement = ".")
  #names(data) <- gsub(x = names(data), pattern = "_cpxii_r1_hs.", replacement = ".")
  names(data) <- gsub(x = names(data), pattern = "_ct_r2_hsms.", replacement = ".")
  
  # Summarize the images and total artifacts
  totals <- colMax(data[names(data)[grep(".Found$", names(data))]] +
           data[names(data)[grep(".Remaining$", names(data))]])
  names(totals) <- gsub(x = names(totals), pattern = "Found", replacement = "Total") 
  names(totals) <- unlist(lapply(strsplit(names(totals), "_"), `[[`, 1))
  
  
  imagetotals <- colSums(!is.na(data[names(data)[grep(".Found$", names(data))]]))
  names(imagetotals) <- gsub(x = names(imagetotals), pattern = ".Found", replacement = "") 
  
  # Add Known team names from lookup file
  lookup <- read.csv("lookup.csv", sep=",", header = TRUE, stringsAsFactor = TRUE)
  data <- merge(lookup, data, by = "TeamNumber", all=TRUE, incomparables = '')
  data$TeamName <- NULL
  names(data)[names(data) == 'Name'] <- 'TeamName'
  data$TeamName <- as.factor(data$TeamName)
  excludes <- data[is.na(data$line),]
  data <- data[!is.na(data$line),]
  
  # Formats all of the "<image>.Time" columns into minutes
  data[names(data)[grep(".Time", names(data))]] <- sapply(data[names(data)[grep(".Time", names(data))]], as.difftime, format = "%H:%M", units = "mins")
  
  # Coerce all of the "<image>.Warn" vectors to factors
  data[names(data)[grep(".Warn", names(data))]] <- lapply(data[names(data)[grep(".Warn", names(data))]], as.factor)
  
  # Isolate the Middle School division (original preserved for comparison against the overall population)
  # originaldata <- data
  # data <- data[ which(Division=='Middle School'),]
  
  # Add the completion ratios to each team's scores
  # [TODO] Change the field names to match the competition images
  data$Server2016.Completion <- data$Server2016.Found/totals[1]
  data$Ubuntu14.Completion <- data$Ubuntu14.Found/totals[2]
  data$Windows10.Completion <- data$Windows10.Found/totals[3]
  
  data$State <- as.factor(data$State)
  data$StatePlace <- as.integer(data$StatePlace)
  data$TeamNumber <- as.factor(data$TeamNumber)
  data$Tier <- as.factor(data$Tier)
  data$CurrentScore <- as.integer(data$CurrentScore)
  
  row.names(data) <- NULL
  
  
  # Remove unused columns (i.e. in this particular analysis)
  data$Tier <- NULL
  # data$Division <- NULL
  data$X.1 <- NULL
  
  data <- data[order(data$OverallPlace),]
  top50 <- droplevels(data[data$OverallPlace >= 1 & data$OverallPlace <= 25,])
  ga <- droplevels(data[data$State == "GA",])
  cc <- droplevels(data[!is.na(data$TeamName),])
  top50 <- top50[order(top50$OverallPlace),]
  ga <- ga[order(ga$StatePlace),]
  cc <- ga[order(cc$StatePlace),]
  ga$Difference <- as.integer(c(0, diff(ga$CurrentScore)))
  topscore <- ga$CurrentScore[1]
  ga$OffLead <- ga$CurrentScore-topscore
  
  nteams <- nrow(data)
  myteams <- droplevels(data[data$TeamNumber %in% teams,])
#  data$CurrentScore[is.na(data$CurrentScore)] <- 0

```

# INTRODUCTION
The National Youth Cyber Defense Competition is the centerpiece of the Cyberpatriot program. 
Teams compete by finding cybersecurity vulnerabilities on a set of VMWare virtual machines, then fixing them while maintaining functionality of critical services. 
Teams compete during a six hour period for the top placement within their state and region. 
The first round of the 2019 competition season took place on October 25-27, 2019.
This, the second round took place November 15-17. 
The last competition in the regular season will conclude with the State Round on December 6-8.

<!-- Team,Name -->
<!-- 12-1306,Senior Scripters -->
<!-- 12-1307,Dark_Knights.exe -->
<!-- 12-1308,Cyber.Bytes_Intellect -->
<!-- 12-1309,Code Red -->

Evans Middle School started competing last year in 2017 (CP-10) with two five-student teams. 
The second year of competing expanded to three teams; two with six students, and one with five students. 
This year, there are four teams: three with four students and one with five students. 
The Evans Middle School teams are the *````Senior Scripters````* (```12-1306```), *```Dark_Knights.exe```* (```12-1307```), *````Cyber.Bytes_Intellect````* (```12-1308```),  and *```Code Red```* (```12-1309```).

> [TODO] Participation in the county is increasing as well.

All four Evans Middle School teams competed in this round against a total of ```r nrow(data)``` teams in the nation, including ```r nrow(ga)``` in Georgia. 
Note that ```r nrow(cc)``` of the Georgia schools represent local competition from other Columbia County schools.
Five of the Middle School teams in Georgia were competitive for the National Finals Competition, placing in the overall top 25. 
Evans Middle School placed ```r ordinal(myteams$StatePlace)[1]```, ```r ordinal(myteams$StatePlace)[2]```, and ```r ordinal(myteams$StatePlace)[3]``` among the Georgia teams. 
They were in the top 25% nationally, placing ```r ordinal(myteams$StatePlace)[1]```, ```r ordinal(myteams$StatePlace)[2]```, and ```r ordinal(myteams$StatePlace)[3]```.

The following analysis compares the Evans Middle School teams' performance among the national, state, and local results from the second round in order to develop strategic insights for subsequent competitions. 
The insights can influence the coach and mentors to focus training materials in a way that increases the teams' competitiveness.

# COMPETITION DESCRIPTION 

During each round, the unofficial scores are delivered in near-real time via the competition website (```http://scoreboard.uscyberpatriot.org/```). 
The main page displays a table of key competition data, such as current score and elapsed time. 
Each row on the table contains a secondary link that provides supplementary data, such as a the number of artifacts found, artifacts remaining, and score per image as well as details on any penalties. 
The analysis presented in this document flattens the data in order to extract potential insights that would otherwise remain unknown.

There were ```r length(imagetotals)``` images in this round of competion. The table below shows the total number of teams that attempted each image.

```{r imagetotals, tidy=TRUE}
  print(imagetotals)
```

Each image contained a number of artifacts for a total of 100 possible points for each image.

```{r imagesummary, tidy=TRUE}
  print(totals)
```

A majority of the competing teams used the `cpxii_r2_hsms` images for Windows 10, Server 2016, and Ubuntu 14. 
There were ```r nrow(data[grep("A",data$warn),])``` teams that competed using alternate images for Windows 10 and Server 2016. 
These alternate images were labeled with a `ct_r2_hsms` designation. 
All of the teams using the alternate images were from Canada, though most of the ``r length(data$State[data$State=="CAN"])`` Canadian teams used the `cpxii_r2_hsms` images. 
The scoring for the alternate images did not significantly affect the outcome of the contest.

```{r alternate summary, tidy=TRUE}
alts <- data[grep("A",data$warn),c(1, 16, 7, 5, 30:32, 4)]
alts[is.na(alts)] <- 0
alts$Server2016.Completion <- percent(alts$Server2016.Completion)
alts$Ubuntu14.Completion <- percent(alts$Ubuntu14.Completion)
alts$Windows10.Completion <- percent(alts$Windows10.Completion)
alts
```
# SCORING ANALYSIS 

The overall results from the Evans Middle School teams are in the table below.
_Note that the units for ```PlayTime``` fields are elapsed minutes since the image was initialized._ 
_Teams have six continuous hours (i.e., 360 minutes) and must officially close their image before the deadline to avoid a penalty._

## Team Results

```{r myteamresults}
  vars <- c("TeamName", "StatePlace", "OverallPlace", "CurrentScore", "PlayTime")
  print(myteams[vars], row.names = FALSE)
```

*````Senior Scripters````* and *```Dark_Knights.exe```* tied for completion on the Ubuntu 14 image.
*````Senior Scripters````* found two more artifacts over *```Dark_Knights.exe```* on the Windows 10 image.
*```Dark_Knights.exe```* had a single artifact advantage on the Server 2016 image.
The result was a separation of four points in the overall score.
Both of the new teams this year--*````Cyber.Bytes_Intellect````* and *```Code Red```*--had strong results on all three images.
The Windows completion rates were slightly lower than Round 1, likely due to the increased difficulty. 
The Ubuntu image was also the first attempt for both of the new teams.

The following tables summarize the team results for each of the images. 
More detailed analysis is included in the _Completion Analysis_ section, later in this document.

### Windows 10 (`r totals[["Windows10.Total"]]` possible artifacts)
```{r myteamWindows10}
  vars <- c("TeamName", "Windows10.Found", "Windows10.Completion", "Windows10.Score", "Windows10.Time")
  disp <- myteams[vars]
  names(disp) <- c("Name", "Found", "Complete(%)", "Score", "Time(min)")
  disp$`Complete(%)` <- percent(disp$`Complete(%)`)
  print(disp, row.names = FALSE)
```

### Server 2016 (`r totals[["Server2016.Total"]]` possible artifacts)
```{r myteamServer2016}
  vars <- c("TeamName", "Server2016.Found", "Server2016.Completion", "Server2016.Score", "Server2016.Time")
  disp <- myteams[vars]
  names(disp) <- c("Name", "Found", "Complete(%)", "Score", "Time(min)")
  disp$`Complete(%)` <- percent(disp$`Complete(%)`)
  print(disp, row.names = FALSE)
```

### Ubuntu 14 (`r totals[["Ubuntu14.Total"]]` possible artifacts)
```{r myteamUbuntu14}
  vars <- c("TeamName", "Ubuntu14.Found", "Ubuntu14.Completion", "Ubuntu14.Score", "Ubuntu14.Time")
  disp <- myteams[vars]
  disp[is.na(disp)] <- 0
  names(disp) <- c("Name", "Found", "Complete(%)", "Score", "Time(min)")
  disp$`Complete(%)` <- percent(disp$`Complete(%)`)
  print(disp, row.names = FALSE)
```

## Competition Results

The analysis in this section presents the overall results of the second round. 
Then, the results of the Evans Middle School teams is compared against the national, state, and the local community results.

### Top `n` Teams
```{r topnsummary}
  # Extract the subsets of interest
  numericInput("rows", "How many Teams?", 25)
  
  renderTable({
    vars <- c("OverallPlace", "TeamNumber", "State", "StatePlace", "CurrentScore")
    head(data[vars], input$rows)
  })
```

### Scores from Georgia schools
Team listed in the table below are registered but did not compete in the semi-finals round.
```{r excludes}
  vars <- c("TeamNumber", "TeamName")
  print(excludes[vars], row.names = FALSE)
```
_NOTE: Teams with a `TeamName` represent Columbia County middle schools with Cyberpatriot teams._ 
```{r statesummary}
  renderTable({
    vars <- c("StatePlace", "TeamNumber", "TeamName", "OverallPlace", "CurrentScore", "Difference", "OffLead", "warn")
    head(ga[vars], nrow(ga))  
  }, row.names = FALSE)
  
```

## Score Comparision

The plots below summarize the unofficial scores by state. 
The numbers along the bottom of the x-axis note the number of teams in each state. 
The horizontal dashed lines represent the national median score (`r median(data$CurrentScore)`) and the first and third quartiles. 
Blue highlights represent Georgia on the left plot, and the Columbia County Middle Schools on the right plot. 
The Evans Middle School teams are highlighted in red on the right plot.

The median score for the state of Georgia (`r median(ga$CurrentScore)`) is higher than the national median. 
However, the state median falls within the national IQR, suggesting that the difference between the state median and the national median is not statistically significant.
Two of the Evans Middle School teams are in the 25th percentile for the state of Georgia, indicating competitive results.
The other two teams are above the state median (50th percentile), also indicating favorable results.

```{r score}
  give.n <- function(x){
  return(c(y = -10, label = length(x))) 
  }

  p1 <- ggplot(data, aes(State, CurrentScore)) +
    geom_boxplot() +
    ggtitle("Overall Scores by State") + labs(x = "State", y = "Score") +
    theme(axis.text.x=element_text(size=6)) + 
    scale_y_continuous(breaks=seq(0,300,20)) +
    stat_summary(fun.data = give.n, geom = "text", size = 2, position = position_dodge(width = 1), fun.y = median) +
    geom_boxplot(data=ga, aes(State, CurrentScore),fill="blue") +
    geom_line() + geom_hline(yintercept = median(data$CurrentScore), lty=2) + 
    geom_line() + geom_hline(yintercept = quantile(data$CurrentScore, c(.25, .75)), lty=3)
    
    
  p2 <- ggplot(data=ga, aes(State, CurrentScore)) +
      geom_boxplot(outlier.size=0) +
      geom_jitter(data=data[data$TeamName != "",], aes(State, CurrentScore),
             position=position_jitter(width=0.1,height=0.1),
             alpha=0.6,
             size=3,
             show.legend = FALSE,
             color = "blue") +
      geom_jitter(data=myteams, aes(State, CurrentScore),
               position=position_jitter(width=0.1,height=0.1),
               alpha=0.6,
               size=3,
               show.legend = FALSE,
               color = "red") +
    scale_y_continuous(breaks=seq(0,300,60)) +
    xlab("Georgia Teams") +
    ylab("Score") +
    ggtitle("GA Teams")
    
    grid.arrange(p1, p2, ncol = 2, top = "Score Analysis", widths = c(5, 1))
    
    dens <- data %>% group_by(State) %>% summarise(nteams = n(), medianscore = median(CurrentScore))
```

### Score Density

Overall placement is determined primarily by overall scores (higher is better).
Numerous teams end up with the same overall score, so time is used to further categorize team placement  (faster is better).
The plots below summarize the Overall Scores and Play Times to provide descriptive context of both variables for the remaining analysis.

```{r densitystats}

## [TODO] Separate into two graphs with different scales
  tm <- data %>% select(OverallPlace, CurrentScore, PlayTime)
  d <- describeBy(tm[c(-1)])
  print(d[,c(-1,-6,-7)], row.names = FALSE)
  
  d <- melt(tm, id.vars="OverallPlace")
  p1 <- ggplot(d, aes(variable, value)) + geom_boxplot() +
    xlab("Variable") +
    ylab("Value") +
    scale_y_continuous(breaks=seq(0,max(data$PlayTime, na.rm = TRUE),60)) +
    ggtitle("Summary of Score and Time Variables")
  p1
  
```

The plots below illustrate the density of scores and time.
The red bars represent the Evans Middle School teams.
The blue bars represent the State of Georgia teams.
The gray bars represent the national/international teams, or the overall field of competition.
The vertical dotted line represents the median score.
The Georgia teams are spread on either side of the median. 
The scores for the Evans Middle School scores were in competitive groups, and consistently above the median.

The median play time was `r median(data$PlayTime, na.rm = TRUE)` minutes.
`r length(data$PlayTime[which(data$PlayTime > 360)])` teams (`r length(ga$PlayTime[which(ga$PlayTime > 360)])` from Georgia) exceeded the 6-hour (i.e., 360 minutes) time limit.
The Evans Middle School Teams closed out the challenges slightly early, though slightly longer than the overall, with a median play time of `r median(myteams$PlayTime)` minutes.

```{r scoretimedensity}
  p1 <- ggplot(data, aes(CurrentScore)) +
      geom_histogram(binwidth = 1) +
      geom_histogram(data=ga, fill="blue", binwidth = 1) + 
      geom_histogram(data=myteams, fill="red", binwidth = 1) + 
      scale_x_continuous(breaks=seq(0,max(data$CurrentScore, na.rm = TRUE),10)) +
      scale_y_continuous(breaks=seq(0,25,2)) +
      geom_vline(xintercept = median(data$CurrentScore, na.rm = TRUE), lty=2) +
      ggtitle("Scoring Density") + labs(x = "Overall Score", y = "Number of Teams")
  
  p1
  
  p2 <- ggplot(data, aes(PlayTime)) +
      geom_histogram(binwidth = 10) + 
      geom_histogram(data=ga, fill="blue", binwidth = 10) + 
      geom_histogram(data=myteams, fill="red", binwidth = 10) + 
      geom_vline(xintercept = median(data$PlayTime, na.rm = TRUE), lty=2) +
      geom_vline(xintercept = 360) +
      scale_x_continuous(breaks = seq(0,max(data$PlayTime, na.rm = TRUE),60)) +
      scale_y_continuous(breaks=seq(0,max(data$CurrentScore, na.rm = TRUE),10)) +
      geom_text(aes(x=360, label="Time Limit", y=120), angle=90, vjust = 1, text=element_text(size=3)) +
      ggtitle("Play Time Density") + labs(x = "Play Time (minutes)", y = "Number of Teams")
  
  p2

```

The plot below displays the number of teams (colored boxes) against both play time (x) and overall score and play time (y). 
The plot confirms that the teams that score above the national median tend to play up to within 30 minutes of the limit. 
An exception is the ```r length(data$PlayTime[data$CurrentScore>290])``` highest scoring teams (i.e., score >290) that completed in ```r min(data$PlayTime[data$CurrentScore>290])``` and ```r max(data$PlayTime[data$CurrentScore>290])``` minutes with a median time of ```r median(data$PlayTime[data$CurrentScore>290])```.
  
```{r scorevstime}
  # Two-dimension density plot
  p <- ggplot(data, aes(PlayTime,CurrentScore)) + 
      stat_bin2d(binwidth=c(15,10)) +                       ## bins are 15 minutes and 10 points 
      scale_fill_gradientn(colors=topo.colors(5)) + 

      geom_hline(yintercept = median(data$CurrentScore), lty=2) + 
      geom_vline(xintercept = 360, color="red") +
  
      # geom_smooth(method = lm) +
      scale_x_continuous(breaks = seq(0,max(data$PlayTime, na.rm = TRUE),60)) +
      scale_y_continuous(breaks=seq(0,max(data$CurrentScore, na.rm = TRUE),10)) + 
      ggtitle("Density of Scores and Play Time") + labs(x = "Play Time (minutes)", y = "Overall Score")
  p

  # Scatter plot with smoothed line
  # p1 <- ggplot(tm, aes(x=PlayTime, y=CurrentScore)) + 
  #   geom_point(shape = 1) + 
  #   geom_vline(xintercept = 360) +
  #   geom_smooth(method = lm) +
  #   scale_x_continuous(breaks = seq(0,max(data$PlayTime, na.rm = TRUE),60)) +
  #   scale_y_continuous(breaks=seq(0,max(data$CurrentScore, na.rm = TRUE),10)) +
  #   ggtitle("Score vs. Time") + labs(x = "Play Time (minutes)", y = "Overall Score")
  # 
  # p1

```

### Impact of State Team Density on Overall Scores

A total of ```r nrow(dens)``` states and countries participated in this round. 
The table of diagnostic statistics below shows that each state has a large range of the number of teams as well as overall scores. 
It may seem intuitive that states with more teams lead to better overall scores, 
However, as shown in the _Score Comparison_ section above, the number of teams in each state compared against that state's median score suggests that there is very little statistical relationship between the two.
The figures below summarize the descriptive statistics and density of teams per state/country.

```{r teamdensitystats}
  d <- describeBy(dens[c(-1)])
  print(d[,c(-1,-6,-7)], row.names = FALSE)
```


```{r stateteams}
  stteams <- data.frame(table(data$State))
  names(stteams) <- c("State", "Teams")
  
  p1 <- tableGrob(head(stteams[order(-stteams$Teams),],10), rows = NULL)
  
  p2 <- ggplot(stteams, aes(x=reorder(State, -Teams), Teams)) +
      geom_bar(stat = "identity") +
      labs(x = "State", y = "Teams") +
      theme(axis.text.x=element_text(size=6))

  grid.arrange(p1, p2, ncol = 2, top = "Team Density by State", widths = c(1, 4))
  
```

The plots below illustrate the states with the highest overall scores tend to have very few teams, while the states with the most teams trend along overall median. 

The boxplot on the left provides the context of overall team density.
The median density for a state is ```r quantile(dens$nteams)[["50%"]]``` teams, with an IQR of ```r quantile(dens$nteams)[["75%"]] - quantile(dens$nteams)[["25%"]]```.
The outliers are plotted vertically.
Georgia, with ```r dens[dens$State=="GA",]$nteams``` teams is one of the outliers.

In contrast, the scatterplot on the right shows each state/country horizontally against their median score.
The points in red show the lowest and highest density states in relation to Georgia.
The blue dashed line represents the overall median of the state median scores (```r quantile(dens$medianscore)[["50%"]]```) with an IQR of ```r quantile(dens$medianscore)[["75%"]] - quantile(dens$medianscore)[["25%"]]```; the IQR is omitted from the scatterplot for clarity.
A small, positive correlation coefficient (```r round(cor(dens$nteams, dens$medianscore), 4)```) suggests that more teams per state do not necessarily translate into higher median scores.
Rather, the plot seems to suggest that states with more teams tend to regress toward the national median score.
The median score for Georgia (```r dens[dens$State=="GA",]$medianscore```) is slightly higher than the overall median score, and within the IQR. 
Therefore, the state of Georgia is competitive but not necessarily due to the higher team density.


```{r teamdensity}
  p1 <- ggplot(dens, aes(x = "", y = nteams)) + 
    geom_boxplot() + 
    labs(x = "", y = "Overall Team Density per State")

  p2 <- ggplot(dens, aes(x = nteams, y = medianscore)) + 
  geom_point(aes(colour = ifelse((dens$State == "GA" | dens$nteams == max(dens$nteams) | dens$medianscore == max(dens$medianscore)), "black", "red")))+
  geom_text(aes(label = ifelse((dens$State == "GA" | dens$nteams == max(dens$nteams) | dens$medianscore == max(dens$medianscore)), paste0(as.character(State), "(", as.character(nteams),")"), "")), hjust = 1, vjust = -0.75, size = 3)+
  theme(legend.position = "none") +
  annotate("text", label = paste0("cor = ", round(cor(dens$nteams, dens$medianscore), 4)), x = 100, y = 0) +
  geom_hline(yintercept = median(data$CurrentScore), color= "blue", lty=2) + 
  labs(x = "Number of teams", y = "Median Score")
  
  grid.arrange(p1, p2, ncol = 2, top = "Team Density Analysis", widths = c(1, 5))
```
_NOTE: Each point represents a State's median score in relation to the number of teams in that state. Therefore, the horizontal blue dashed line represents the median of the overall scores, and not the median of the points represented in the plot (i.e., the median of medians)._

# IMAGE ANALYSIS

## Image Scoring

```{r completionsetup}
  d <- data.frame(table(data$ScoredImages))
  names(d) <- c("Scored Images", "Teams")
```

The Evans Middle School teams attempted all three of the competition images, consistent with ```r percent(d$Teams[3]/nrow(data))``` of the teams in the competition. 
In contrast, ```r d$Teams[1] + d$Teams[2]``` teams or ```r percent((d$Teams[1] + d$Teams[2])/nrow(data))``` (including ```r nrow(droplevels(ga[ga$ScoredImages < 3,]))``` in Georgia) attempted less than three images. 
The teams that did not attempt all of the images ended up placing between ```r min(data[data$ScoredImages < 2,]$OverallPlace)``` and ```r max(data[data$ScoredImages < 3,]$OverallPlace)```.

The plot below shows a large variation among the scores with teams that attempted both images. 
However, the scores from the teams did not attempt all of the images are are lower and statistically significant enough to conclude that they are not competitive.

```{r attempts}

  ggplot(data, aes(x=ScoredImages, y = CurrentScore)) + 
  geom_boxplot(aes(group=ScoredImages)) + 
  geom_jitter(color = "grey50") +
      geom_jitter(data=data[data$TeamName != "",], aes(ScoredImages, CurrentScore),
             position=position_jitter(width=0.05,height=0.1),
             alpha=0.6,
             size=3,
             show.legend = FALSE,
             color = "blue") +
      geom_jitter(data=myteams, aes(ScoredImages, CurrentScore),
               position=position_jitter(width=0.05,height=0.1),
               alpha=0.6,
               size=3,
               show.legend = FALSE,
               color = "red") +
      ggtitle("Scores per Images") + labs(x = "Attempted Images", y = "Score")

```


## Image Completion Rates

The plots below show the ratio of artifacts found compared to number of total artifacts on each image. 
Note that a ratio is used (i.e., instead of number of artifacts found) in order to compare the images of different numbers of available artifacts. 
In this case, the images have different numbers of possible artifacts, so using ratios is preferrable for comparison.
The national trend was that teams completed slightly more than half of the Windows images and almost two-thirds of the Ubuntu image.
As shown in the _Team Results_ section above, all three of the Evans Middle School teams completed more on each image than the nationwide median. 
The one exception was _Code Red_; this was thier first attempt on a competition Ubuntu image, completing 55% of it.

```{r completion}
  d <- data[,c(1, 30:32)]
  d <- melt(data = d, id.vars = "TeamNumber", measure.vars = c("Ubuntu14.Completion", "Windows10.Completion","Server2016.Completion"))
  d$variable <- gsub(".Completion", "", d$variable)
  names(d) <- c("Teams", "Image", "Complete")
  
  db <- describeBy(d[3], group = d$Image, mat = TRUE, digits = 2)
  
  print(db[,c(-1,-3,-8,-9)], row.names = FALSE)
  
  ggplot(d, aes(x=Image, y = Complete)) +
      geom_boxplot() +
    
      geom_jitter(data=d[d$Teams %in% ga$TeamNumber,], aes(Image, Complete),
               #position=position_jitter(width=0.1,height=0.1),
               alpha=0.6,
               size=3,
               show.legend = FALSE,
               color = "blue") +

      geom_jitter(data=d[d$Teams %in% teams,], aes(Image, Complete),
         #position=position_jitter(width=0.1,height=0.1),
         alpha=0.6,
         size=3,
         show.legend = FALSE,
         color = "red") +
    
      ggtitle("Overall Completion Rates per Challenge") + labs(x = "Challenge", y = "% Complete")
```

## Image Artifacts

The plots below count the number of teams that found the same number of artifacts on each image.
The results are consistent with the previosuly stated observations.

```{r imagescore}
  g1 <- ggplot(data, aes(Ubuntu14.Found)) + geom_bar() +
  geom_bar(data=ga, aes(Ubuntu14.Found),fill="blue") +
  geom_bar(data=myteams, aes(Ubuntu14.Found),fill="red", width = 1)

# aes(fill = ifelse(data[data$TeamNumber %in% teams,]$Ubuntu14.Found, "grey50", "red"))

  g2 <- ggplot(data, aes(Windows10.Found)) + geom_bar() + 
    geom_bar(data=ga, aes(Windows10.Found),fill="blue") +
    geom_bar(data=myteams, aes(Windows10.Found),fill="red", width = 1)
  
  g3 <- ggplot(data, aes(Server2016.Found)) + geom_bar() + 
    geom_bar(data=ga, aes(Server2016.Found),fill="blue") +
    geom_bar(data=myteams, aes(Server2016.Found),fill="red", width = 1)
    
  grid.arrange(g1, g2, g3, ncol = 3, top = "Frequency of Finds per Image")

```

# CONCLUSIONS
> [TODO] 
* Compared to the National and State teams, both of the Evans Middle School teams are very competitve on the Windows Server image, but could have been more competitive on the Windows 10 image.
* The number of teams that compete in a given state does not necessarily have a positive effect on overall scores. In fact, the states with between ```7``` and ```12``` teams appear the most competitive.
* Teams that do not attempt both images are not competitive, though attempting both images does not necessarily guarantee high placement.
* [DRAFT] Is there an advantage to finishing earlier?